The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-04-16 paper/iphas-dr2.tex
Finished polishing Sec 4

Despite the best efforts made to obtain a nightly calibration,
large surveys naturally possess field-to-field variations
due to atmospheric changes during the night
and imperfections in the pipeline or the instrument
(e.g. the WFC is known to suffer from sporadic errors
in the timing of exposures).
This is demonstrated in Fig.~\ref{fig:ccd_before},
where we show the combined colour-colour diagram
for all 2,801 quality-approved fields
which are located towards the Galactic anti-centre
($160\degr < \ell < 200\degr$).
The main locus of stars is poorly defined
in the diagram due to the presence of
incorrectly calibrated fields.
The calibration errors can exceed 0.1~mag or more
and need to be corrected for 
during a global re-calibration procedure.
For comparison, Fig.~\ref{fig:ccd_after}
shows the identical diagram which was constructed
after our re-calibration.

\begin{figure}
\captionsetup[subfigure]{labelformat=empty}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/caldiagram/ccd-uncalibrated.pdf}
\caption{}
\label{fig:ccd_before}
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/caldiagram/ccd-calibrated.pdf}
\caption{}
\label{fig:ccd_after}
\end{subfigure}
\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram
         covering an area of 400~deg$^2$,
         shown before (panel a) and after (panel b)
         our global re-calibration was applied.
         Both figures were created by combining stars
         detected across all 2,801 quality-approved fields
         which are located towards the Galactic anti-centre
         (here defined as $160\degr<\ell<200\degr$).
         The diagrams are plotted as 2D-histograms
         which show the density of sources
         in bins of 0.01-by-0.01~mag;
         bins containing 1 to 200 sources are coloured red,
         while bins containing more than 400 sources are bright yellow.
         The diagrams include all stars
         brighter than $r<18$ which were
         classified by the pipeline as being `very reliable'
         (to be explained in
         \S\ref{sec:qualitycriteria}).}
\end{figure}

Notable past examples of surveys which required
global re-calibration include 
2MASS \citep[][]{Nikolaev2000},
SDSS \citep[][]{Padmanabhan2008}
and the Panoramic Survey Telescope 
And Rapid Response System survey \citep[Pan-STARRS;][]{Schlafly2012},
which all achieved photometry 
that is globally consistent to within 0.01--0.02~mag.

Surveys which observe identical stars at different epochs
can use the repeat measurements to ensure a homogeneous calibration.
For example, 2MASS attained its global calibration
by observing six standard fields each hour, 
allowing zeropoint variations to be tracked 
over very short timescales \citep{Nikolaev2000}.
Alternatively, the SDSS and PanSTARRS surveys could benefit
from revisiting regions in their footprint to 
carry out a so-called \emph{ubercalibration}\footnote{`Ubercalibration'
refers to the name of the code used to re-calibrate SDSS photometry
using repeat observations. 
It is an anglicised version of the German word `\"uberkalibration',
which was reportedly chosen because the initial authors both
had German-sounding names \citep{Finkbeiner2010}.} procedure,
in which multiple measurements of stars at different epochs
are used to fit the calibration parameters
\citep[e.g.][]{Ivezic2007,Padmanabhan2008,Schlafly2012}.

@@

We have found the information contained
in these overlap regions to be insufficient
to constrain the calibration parameters.
This is because photometry at the extreme edges of the WFC
-- where neighbouring fields overlap -- 
is prone to systematics at the level of 1--2~per cent.
The cause of these errors is thought to include 
the use of twilight sky flats in the pipeline,
which are known to be imperfect for calibrating stellar photometry 
due to stray light and vignetting \citep[e.g.][]{Manfroid1995}.
Moreover, the illumination correction in the overlap regions
is more affected by a radial geometric distortion in the WFC,
which causes the pixel scale to increase as the edges 
are approached \citep{Gonzalez-Solares2011}.
Although these systematics are reasonably small within a single field,
they can combine, during the global calibration routine, to cause artificial zeropoint gradients 
across the survey unless controlled by other external constraints.

@@

We have been able to benefit from the
AAVSO Photometric All-Sky Survey
(APASS; http://www.aavso.org/apass)
to bring most of the survey 
onto a uniform scale.
Since 2009,
APASS has been using two 20~cm-astrographs
to survey the entire sky down to $\sim$17th magnitude
in five filters which include Sloan $r$ and $i$ \citep{Henden2012}.
The most recent catalogue available 
at the time of preparing this work was APASS DR7,
which provides a good coverage across $\sim$half
of the IPHAS footprint.
The overlap regions are shown in Fig.~\ref{fig:apass_r}.
The photometric accuracy of APASS is currently estimated 
to be at the level of 3 per cent (Henden, private communication),
which is significantly better 
than the original nightly calibrations of IPHAS
which are accurate to $\sim$10 per cent (Tab.~\ref{tbl:offsets_before}).
APASS achieves its uniform accuracy 
by measuring each star at least two times in photometric conditions,
along with ample standard fields,
benefiting from the large 3-by-3 degrees field of view of its detectors.

@@

The resulting set of $\sim$220\,000 cross-matched stars were then used 
to derive APASS-to-IPHAS magnitude transformations
using a linear least-squares fitting routine, 
which iteratively removed $3\sigma$-outliers to improve the fit.
The solution converged to:
\begin{align} 
r_{\rm IPHAS} = r_{\rm APASS} - 0.121 + 0.032(r-i)_{\rm APASS} \label{eqn:apass_r} \\
i_{\rm IPHAS} = i_{\rm APASS} - 0.364 + 0.006(r-i)_{\rm APASS} \label{eqn:apass_i}
\end{align}
The root mean square (rms) residuals of these transformations 
are 0.041 and 0.051, respectively.
The small colour terms in the equations
indicate that the IPHAS and APASS broadband filters 
are very similar.
The transformations include a large fixed offset,
but this is simply due to the fact that 
APASS magnitudes are given in the AB system
and IPHAS uses Vega-based magnitudes.
Separate transformations were derived for sight-lines 
with varying extinction properties to investigate the robustness
of the transformations with respect to different reddening regimes.
The variations at these different
sight-lines were found to be insignificant.
This is not surprising because heavily-reddened objects 
are naturally scarce at $r<16$.

Having transformed APASS magnitudes into the IPHAS system,
we then computed the median magnitude offset 
for each field which contained at least 30 cross-matched stars.
This was achieved for 48 per cent of our fields (shown in Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}).
The offsets follow a near-Gaussian distribution
centred around $0.014\pm0.104$~mag in $r$ and $0.007\pm0.108$~mag in $i$
(Table~\ref{tbl:offsets_before}).
A total of 4596 fields showed a systematic offset
exceeding $\pm$0.03~mag in either $r$ or $i$.

@@

We then applied the most important step in our calibration scheme,
which is to adjust the provisional zeropoints of these 4596 aberrant fields
such that their offset is brought to zero.
This allowed the mean IPHAS-to-APASS offset 
to be brought down to $0.000\pm0.011$~mag in both $r$ and $i$
(Table~\ref{tbl:offsets_after}).
The procedure of fitting magnitude transformations and
correcting the IPHAS zeropoints was repeated a few times to ensure 
convergence, which was essentially reached after the first iteration.

@@

At the time of writing, the APASS catalogue did not provide 
sufficient coverage for 7359 of our fields.
Fortunately, these fields are located mainly 
at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
which were typically observed during the summer months
when photometric conditions are more prevalent at the telescope.
These remaining fields have nevertheless 
been brought onto the same uniform scale 
by employing an ubercalibration-style scheme,
which minimises the magnitude offsets between stars
located in the overlap regions between neighbouring fields.

An algorithm for achieving this minimisation
has previously been described by~\citet{Glazebrook1994}.
In brief, there are two fundamental quantities to be
minimised between each pair of overlapping exposures, 
denoted by the indices $i$ and $j$. 
Firstly, the mean magnitude difference between stars in the overlap
region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
constraint. 
Secondly, to ensure the solution does not stray far 
from the existing calibration, 
the difference in zeropoints 
$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
between each pair of exposures must also be minimised.

Minimisation of these two quantities is a linear least squares problem 
because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
Hence we can find the ZP shift to be applied to each field 
by minimising the sum:
\begin{equation}
   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
   \label{eqn:chi2}
\end{equation}
where $i$ denotes an exposure, 
$j$ an overlapping exposure, 
$N$ the number of exposures,
$a_i$ the ZP to solve for 
and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
and $\theta_{ij}$ is an overlap function 
equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
Solving for $a_i$ is equivalent to solving $\partial
S/\partial a_i=0$, which gives the matrix equation:
\begin{equation}
   \sum_{j=1}^N A_{ij} a_j = b_j,
   \label{eqn:matrix}
\end{equation}
where 
\begin{eqnarray}
   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
\end{eqnarray}

We enforce a strong external constraint
on the solution by keeping the zeropoint fixed 
for the 
% 6\,756 
fields which have already been compared
and calibrated against APASS.
We hereafter refer to these fields as \emph{anchors}.
It is asserted that the zeropoints $a_i$ of the anchor fields 
are known and not solved for.  However they do appear in the vector 
$b_j$ as constraints.  In addition to the APASS-based anchors, 
we selected 3\,273 additional anchor fields by hand
to provide additional constraints in regions not covered by APASS.
These extra anchors were deemed to have reliable zeropoints 
based on 
(i) the information contained in the observing logs,
(ii) the stability of the standard star zeropoints during the night, and
(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
which is located $\sim$500~m from the INT.

We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
separately using the least-squares routine 
in Python's {\sc scipy.sparse} module for sparse matrix algebra.
This provided us with corrected zeropoints for the remaining fields,
which were shifted on average by $+0.02\pm0.11$ in $r$ 
and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.
The outcome is validated in the next section.

We then turned to the global calibration of the \ha\ data.
It is not possible to re-calibrate the narrowband 
in the same way as the broadbands,
because the APASS survey does not offer \ha\ photometry.
We can reasonably assume however,
that the corrections required for $r$ and \ha\ are identical,
much of the time, because the IPHAS data-taking pattern ensured 
that a field's \ha\ and $r$-band exposures
were taken consecutively,
albeit separated by the $\sim$30~s read-out time.
Hence, we have corrected the \ha\ zeropoints 
by re-using the zeropoint adjustments that were derived for the $r$ band
in the earlier steps.
An exception was made for a small subset of fields
% 3\,101 fields
for which our quality-control routines revealed
strong zeropoint variations during the night,
suggesting non-photometric conditions.
In these cases, the \ha\ zeropoints
were adjusted by solving Eqn.~\ref{eqn:matrix}
rather than being linked to the $r$-band shift.

@@

Having re-calibrated all fields to the expected APASS accuracy of 3 per cent,
we turned to a different survey, SDSS Data Release 9 \citep{Ahn2012},
to validate the results.
SDSS DR9 includes several strips at low
Galactic latitudes,
providing data across 18 per cent of the fields in our data release.
We cross-matched the IPHAS fields against the subset of
objects marked as reliable stars in the SDSS catalogue\footnote{
We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
to obtain photometry from the SDSS {\sc photoprimary} table 
with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}
cross-matching in much the same way as for APASS,
with the difference of selecting from the  fainter magnitude ranges of 
$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
This provided us with a set of 1.2~million cross-matched stars.

@@

Having transformed SDSS magnitudes into the IPHAS system,
we then computed the median magnitude offset for each IPHAS field
which contained at least 30 objects with a cross-matched counterpart
in SDSS.  This was the case for 2\,602 fields.
The median offsets for each of these fields
are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
Importantly, the mean offset and standard deviation found 
is $-0.001\pm0.029$~mag in $r$
and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
In comparison, offsets computed in the identical way
\emph{before} carrying out our re-calibration showed means
of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
We conclude that our re-calibration procedure has
been successful in improving the
uniformity of the calibration by a factor three
and has achieved our aim of bringing the
accuracy down to the level of 0.03~mag (rms).

@@

In previous work, we have shown that the majority of
\ha-emitters seen by IPHAS towards an H{\sc ii} region
are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
These are young objects
which are thought to show \ha\ in emission
due to the presence of hot, infalling gas
which is accreting onto the star
from a circumstellar disk.
This is likely to be the case for the candidate \ha-emitters
we discovered towards Sh 2-82 as well.
Two of our candidates, \#1 and \#4 in Tab.~\ref{tbl:emitters},
have previously been identified
as candidate YSOs by \cite{Robitaille2008}
and \cite{Yu2012}, respectively.
In these studies, the authors used Spitzer data
to find intrinsically red objects,
which is consistent with the presence of a circumstellar disk.
Although the other four candidate emitters in our sample
have not previously appeared in the literature,
we note that all four are detected
in the Spitzer 8.0~\micron\ image at S/N$>$5.
They are likely to be YSOs exhibiting a mild infrared excess.

>>>>

Despite the best efforts made to obtain a nightly calibration,
large surveys naturally possess field-to-field variations
due to atmospheric changes during the night
and imperfections in the pipeline or the instrument
(e.g. the WFC is known to suffer from sporadic errors
in the timing of exposures).
This is demonstrated in Fig.~\ref{fig:ccd_before},
where we show the combined colour-colour diagram
for all 2,801 quality-approved fields
which are located towards the Galactic anti-centre
(defined here as $160\degr < \ell < 200\degr$).
The main locus of stars is poorly defined
in the diagram due to the presence of
incorrectly calibrated fields.
The calibration errors can exceed 0.1~mag or more
and need to be corrected for 
during a global re-calibration procedure.
For comparison, Fig.~\ref{fig:ccd_after}
shows the identical diagram which was constructed
after our re-calibration procedure.

\begin{figure}
\captionsetup[subfigure]{labelformat=empty}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/caldiagram/ccd-uncalibrated.pdf}
\caption{}
\label{fig:ccd_before}
\end{subfigure}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=\textwidth]{figures/caldiagram/ccd-calibrated.pdf}
\caption{}
\label{fig:ccd_after}
\end{subfigure}
\caption{IPHAS $(r-\ha,\ r-i)$ colour-colour diagram
         covering an area of 400~deg$^2$,
         shown before (panel a) and after (panel b)
         our global re-calibration was applied.
         Both figures were created by combining stars
         detected across all 2,801 quality-approved fields
         which are located towards the Galactic anti-centre
         ($160\degr<\ell<200\degr$).
         The diagrams are plotted as 2D-histograms
         which show the density of sources
         in bins of 0.01-by-0.01~mag;
         bins containing 1 to 200 sources are coloured red,
         while bins containing more than 400 sources are bright yellow.
         The diagrams include all stars
         brighter than $r<18$ which were
         classified by the pipeline as being `very reliable'
         (to be explained in
         \S\ref{sec:qualitycriteria}).
         The objects seen to fall above the locus of stars
         \emph{after} re-calibration 
         are likely to be genuine \ha\ emission-line objects.}
\end{figure}

Notable past examples of surveys which required
global re-calibration include 
2MASS \citep[][]{Nikolaev2000},
SDSS \citep[][]{Padmanabhan2008}
and the Panoramic Survey Telescope 
And Rapid Response System survey \citep[Pan-STARRS;][]{Schlafly2012},
which all achieved photometry 
that is globally consistent to within 0.01--0.02~mag.
Surveys which observe identical stars at different epochs
can use the repeat measurements to ensure a homogeneous calibration.
For example, 2MASS attained its global calibration
by observing six standard fields each hour, 
allowing zeropoint variations to be tracked 
over very short timescales \citep{Nikolaev2000}.
Alternatively, the SDSS and PanSTARRS surveys could benefit
from revisiting regions in their footprint to 
carry out a so-called \emph{ubercalibration}\footnote{`Ubercalibration'
refers to the name of the code used to re-calibrate SDSS photometry
using repeat observations. 
It is an anglicised version of the German word `\"uberkalibration',
which was reportedly chosen because the initial authors both
had German-sounding names \citep{Finkbeiner2010}.} procedure,
in which multiple measurements of stars at different epochs
are used to fit the calibration parameters
\citep{Ivezic2007,Padmanabhan2008,Schlafly2012}.

@@

We have found the information contained
in these narrow overlap regions to be insufficient
to constrain the calibration parameters.
This is because photometry at the extreme edges of the WFC
-- where neighbouring fields overlap -- 
is prone to systematics at the level of 1--2~per cent.
The cause of these errors is thought to include 
the use of twilight sky flats in the pipeline,
which are known to be imperfect for calibrating stellar photometry 
due to stray light and vignetting \citep[e.g.][]{Manfroid1995}.
Moreover, the illumination correction in the overlap regions
is more affected by a radial geometric distortion in the WFC,
which causes the pixel scale to increase as the edges 
are approached \citep{Gonzalez-Solares2011}.
Although these systematics are reasonably small within a single field,
they can combine, during the global calibration routine, to cause artificial zeropoint gradients 
across the survey unless controlled by other external constraints.

@@

We have been able to benefit from the
AAVSO Photometric All-Sky Survey
(APASS; http://www.aavso.org/apass)
to bring most of the survey 
onto a uniform scale.
Since 2009,
APASS has been using two 20~cm-astrographs
to survey the entire sky down to $\sim$17th magnitude
in five filters which include Sloan $r$ and $i$ \citep{Henden2012}.
The most recent catalogue available 
at the time of preparing this work was APASS DR7,
which provides a good coverage across $\sim$half
of the IPHAS footprint.
The overlap regions are shown in Fig.~\ref{fig:apass_r}.
The photometric accuracy of APASS is currently estimated 
to be at the level of 3 per cent (Henden, private communication),
which is significantly better 
than the original nightly calibrations of IPHAS
which are only accurate to $\sim$10 per cent
when compared to APASS (Table~\ref{tbl:offsets_before}).
APASS achieves its uniform accuracy 
by measuring each star at least two times in photometric conditions,
along with ample standard fields,
benefiting from the large 3-by-3 degrees field of view of its detectors.

@@

The resulting set of 220,000 cross-matched stars were then used 
to derive APASS-to-IPHAS magnitude transformations
using a linear least-squares fitting routine, 
which iteratively removed $3\sigma$-outliers to improve the fit.
The solution converged to:
\begin{align} 
r_{\rm IPHAS} = r_{\rm APASS} - 0.121 + 0.032(r-i)_{\rm APASS} \label{eqn:apass_r} \\
i_{\rm IPHAS} = i_{\rm APASS} - 0.364 + 0.006(r-i)_{\rm APASS} \label{eqn:apass_i}
\end{align}
The Root Mean Square (rms) residuals of these transformations 
are 0.041 and 0.051, respectively.
The small colour terms in the equations
indicate that the IPHAS and APASS broadband filters 
are very similar.
The transformations include a large fixed offset,
but this is simply due to the fact that 
APASS magnitudes are given in the AB system
and IPHAS uses Vega-based magnitudes.
Separate transformations were derived for sight-lines 
with varying extinction properties to investigate the robustness
of the transformations with respect to different reddening regimes.
The variations at these different
sight-lines were found to be insignificant.
This is not surprising because heavily-reddened objects 
are naturally scarce at $r<16$.

Having transformed APASS magnitudes into the IPHAS system,
we then computed the median magnitude offset 
for each field which contained at least 30 cross-matched stars.
This was achieved for 48 per cent of our fields (shown in Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}).
The offsets follow a near-Gaussian distribution
with mean and sigma $0.014\pm0.104$~mag in $r$
and $0.007\pm0.108$~mag in $i$
(Table~\ref{tbl:offsets_before}).
A total of 4,596 IPHAS fields showed a systematic offset
exceeding $\pm$0.03~mag in either $r$ or $i$
when compared to APASS.

@@

We then applied the most important step
in our re-calibration scheme,
which is to adjust the provisional zeropoints
of these 4,596 aberrant fields
such that their offset is brought to zero.
This allowed the mean IPHAS-to-APASS offset 
to be brought down to $0.000\pm0.011$~mag in both $r$ and $i$
(Table~\ref{tbl:offsets_after}).
The procedure of fitting magnitude transformations and
correcting the IPHAS zeropoints was repeated a few times to ensure 
convergence, which was essentially reached after the first iteration.

@@

At the time of writing, the APASS catalogue did not provide 
sufficient coverage for 7,359 of our fields.
Fortunately, these fields are located mainly 
at low Galactic longitudes (cf. Figs.~\ref{fig:apass_r} and \ref{fig:apass_i}),
which were typically observed during the summer months
when photometric conditions are more prevalent at the telescope.
These remaining fields have nevertheless 
been brought onto the same uniform scale 
by employing an ubercalibration-style scheme,
which minimises the magnitude offsets between stars
located in the overlap regions between neighbouring fields.

An algorithm for achieving this minimisation
has previously been described by~\citet{Glazebrook1994}.
In brief, there are two fundamental quantities to be
minimised between each pair of overlapping exposures, 
denoted by the indices $i$ and $j$. 
First, the mean magnitude difference between stars in the overlap
region $\Delta_{ij}=\langle m_i-m_j\rangle=-\Delta_{ji}$ is a local
constraint. 
Second, to ensure the solution does not stray far 
from the existing calibration, 
the difference in zeropoints 
$\Delta\mathrm{ZP}_{ij}=-\Delta\mathrm{ZP}_{ji}$ 
between each pair of exposures must also be minimised.

Minimisation of these two quantities is a linear least squares problem 
because the magnitude $m$ depends linearly on the ZP (Eqn.~\ref{eqn:mag}).
Hence we can find the ZP shift to be applied to each field 
by minimising the sum:
\begin{equation}
   S = \sum_{i=1}^N \sum_{j=1}^N w_{ij} \theta_{ij} (\Delta_{ij} + a_i - a_j)^2,
   \label{eqn:chi2}
\end{equation}
where $i$ denotes an exposure, 
$j$ an overlapping exposure, 
$N$ the number of exposures,
$a_i$ the ZP to solve for, 
and $a_j$ the ZP of an overlapping field ($\Delta\mathrm{ZP}_{ij}=a_i-a_j$). 
$w_{ij}$ are weights set to the inverse square of the uncertainty in $\Delta_{ij}$,
and $\theta_{ij}$ is an overlap function 
equal to either 1 if exposures $i$ and $j$ overlap or 0 otherwise. 
Solving for $a_i$ is equivalent to solving $\partial
S/\partial a_i=0$, which gives the matrix equation:
\begin{equation}
   \sum_{j=1}^N A_{ij} a_j = b_j,
   \label{eqn:matrix}
\end{equation}
where 
\begin{eqnarray}
   A_{ij} &=& \delta_{ij} \sum_{k=1}^N w_{jk}\theta_{jk} - w_{ij} \theta_{ij},\label{eqn:aij}\\
   b_i &=& \sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ji} = -\sum_{j=1}^N w_{ij} \theta_{ij}\Delta_{ij}.\label{eqn:bi}
\end{eqnarray}

We enforce a strong external constraint
on the solution by keeping the zeropoint fixed 
for the 
% 6\,756 
fields which have already been compared
and calibrated against APASS.
We hereafter refer to these fields as \emph{anchors}.
It is asserted that the zeropoints $a_i$ of the anchor fields 
are known and not solved for.  However they do appear in the vector 
$b_j$ as constraints.  In addition to the APASS-based anchors, 
we selected 3,273 additional anchor fields by hand
to provide additional constraints in regions not covered by APASS.
These extra anchors were deemed to have reliable zeropoints 
based on 
(i) the information contained in the observing logs,
(ii) the stability of the standard star zeropoints during the night, and
(iii) photometricity statistics provided by the Carlsberg Meridian Telescope,
which is located $\sim$500~m from the INT.

We then solved Eqn.~\ref{eqn:matrix} for the $r$ and $i$ bands
separately using the least-squares routine 
in Python's {\sc scipy.sparse} module for sparse matrix algebra.
This provided us with corrected zeropoints for the remaining fields,
which were shifted on average by $+0.02\pm0.11$ in $r$ 
and $+0.01\pm0.12$ in $i$ compared to their provisional calibration.

We then turned to the global calibration of the \ha\ data.
It is not possible to re-calibrate the narrowband 
in the same way as the broadbands,
because the APASS survey does not offer \ha\ photometry.
We can reasonably assume, however,
that the corrections required for $r$ and \ha\ are identical,
much of the time, because the IPHAS data-taking pattern ensured 
that a field's \ha\ and $r$-band exposures
were taken consecutively,
albeit separated by a $\sim$30~s read-out time.
Hence, we have corrected the \ha\ zeropoints 
by re-using the zeropoint adjustments that were derived for the $r$ band
in the earlier steps.
An exception was made for 3,101 fields
for which our quality-control routines revealed
strong zeropoint variations to exist during the night,
suggesting non-photometric conditions.
In these cases, the \ha\ zeropoints
were adjusted by solving Eqn.~\ref{eqn:matrix}
rather than being linked to the $r$-band shift.

@@

Having re-calibrated all fields
to the expected APASS accuracy of 3 per cent,
we turned to a different survey,
SDSS Data Release 9 \citep{Ahn2012},
to validate the results.
SDSS DR9 includes several strips at low
Galactic latitudes,
providing data across 18 per cent of the fields in our data release.
We cross-matched the IPHAS fields against the subset of
objects marked as reliable stars in the SDSS catalogue\footnote{
We used the CasJobs facility located at http://skyserver.sdss3.org/CasJobs
to obtain photometry from the SDSS {\sc photoprimary} table 
with criteria {\sc type = star}, {\sc clean = 1} and {\sc score $>$ 0.7}.}, in much the same way as for APASS,
with the difference of selecting from the  fainter magnitude ranges of 
$15<r_{\rm SDSS}<18.0$ and $14.5<i_{\rm SDSS}<17.5$.
This provided us with a set of 1.2~million cross-matched stars.

@@

Having transformed SDSS magnitudes into the IPHAS system,
we then computed the median magnitude offset for each IPHAS field
which contained at least 30 objects with a cross-matched counterpart
in SDSS.  This was the case for 2,602 fields.
The median offsets for each of these fields
are shown in Figs.~\ref{fig:sdss_r} and \ref{fig:sdss_i}.
Importantly, the mean offset and standard deviation found 
is $-0.001\pm0.029$~mag in $r$
and $-0.002\pm0.032$~mag in $i$ (Table~\ref{tbl:offsets_after}).
In comparison, offsets computed in the identical way
\emph{before} carrying out the re-calibration showed means
of $+0.016\pm0.088$~mag in $r$ and $+0.010\pm0.089$~mag in $i$ (Table~\ref{tbl:offsets_after}).
We conclude that our re-calibration procedure has
been successful in improving the
uniformity of the calibration
by a factor three (i.e. from $\sigma=0.088$ to $\sigma=0.029$ in $r$),
and as such has achieved our aim of bringing the
accuracy to the aimed level of 0.03~mag.

@@

In previous work, we have shown that the majority of
\ha-emitters seen by IPHAS towards an H{\sc ii} region
are likely to be Classical T Tauri Stars \citep{Barentsen2011a}.
These are young objects
which are thought to show \ha\ in emission
due to the presence of hot, infalling gas
which is accreting onto the star
from a circumstellar disk.
This is likely to be the case for the candidate \ha-emitters
we discovered towards Sh 2-82 as well.
Two of our candidates, \#1 and \#4 in Table~\ref{tbl:emitters},
have previously been identified
as candidate YSOs by \cite{Robitaille2008}
and \cite{Yu2012}, respectively.
In these studies, the authors used Spitzer data
to find intrinsically red objects,
which is consistent with the presence of a circumstellar disk.
Although the other four candidate emitters in our sample
have not previously appeared in the literature,
we note that all four are detected
in the Spitzer 8.0~\micron\ image at S/N$>$5.
They are likely to be YSOs exhibiting a mild infrared excess.
